Global Optimization of an Analog Method by Means of Genetic Algorithms

### 2016-09-14 Genetic algorithms - MWR/template.tex
Edition following AMS guidelines.

% TODO: AMS guidelines
% TODO: compléter DOI

@@

\abstract{Analog methods are based on a statistical relationship between synoptic meteorological variables (predictors) and local weather (predictand, which are to be predicted), which is defined by several parameters that are often calibrated by means of a semi-automatic sequential procedure. This calibration approach is fast, but has strong limitations: it is made of successive steps and thus cannot handle all parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows (hour(s) of the day) on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely Genetic Algorithms, which can optimize jointly all parameters of the method and can get closer to a global optimum, by taking into account the parameters dependencies. Moreover, it can objectively calibrate parameters that were previously manually assessed, and can take into account new degrees of freedom.
Genetic Algorithms need however to be tailored to the addressed problem. Multiple combinations of algorithms were assessed, and new ones were developed (e.g. the \textit{chromosome of adaptive search radius} which was found to be very robust), in order to make recommendations for the use of Genetic Algorithms to optimize several variants of analog methods. A global optimization approach opens new perspectives for the improvement of analog methods, and their application to new regions or to new predictands.
}

@@

Analog methods (AMs) rely on the hypothesis that similar situations in terms of the atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. They consist in sampling a certain number of past situations based on different synoptic-scale meteorological variables (predictors) in order to build a probabilistic prediction for a local weather variable of interest (predictand). A common usage of AMs is for operational precipitation forecasting \citep[eg.][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015b, BenDaoud2016}, or more recently for precipitation downscaling in a climate perspective \citep[eg.][]{Radanovics2013, Chardon2014, Dayon2015}. However, AMs or equivalent are also used to predict temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

@@

A common approach to optimize the method is by means of a semi-automatic sequential calibration procedure, developed by \citet{Bontron2004}, also described in \citet{BenDaoud2016}, and extended by \citet{Radanovics2013}. It determines some parameters of the method, sequentially for each consecutive analogy level (eg. on the atmospheric circulation or on a moisture index). It starts by a manual selection of the meteorological variables (e.g. geopotential height, relative humidity, etc.), the pressure levels, the temporal windows, and the initial analog numbers. Then, the spatial window on which the predictors are compared is optimized by an iterative growth of the domain, and the number of analogs is finally reassessed. A successive level of analogy can then be added and its spatial window optimized. Parameters of the preceding levels of analogy are not reassessed, except for the number of analog situations to keep.

@@

When creating the sequential calibration procedure, \citet{Bontron2004} was aware of the problem of dependencies between parameters and wrote: ''\textit{We perceive here the combinatorial aspect of our problem: variables and spatial windows are not independent. We will present our results by first searching the best variable [note: e.g. selection of the pressure level and the temporal window for the geopotential height] on a chosen spatial window, and next, the best window for the chosen variable. However, even by repeating the process, are we sure to obtain the optimal combination?}''. And later in his work: ''\textit{Our approach, which is again to vary the parameters one by one -- the others being fixed in a more or less arbitrary manner -- may therefore not exactly lead us to the optimal solution}''. \citet{Bliefernicht2010} has also faced the combinatorial issue of AMs parameters and concluded that one needs to be an expert to have a sense of their respective influence, sensitivity and nonlinear interactions. \citet{BenDaoud2010}, when calibrating an AM, also stated that ''\textit{the combinatory aspect related to the calibration was found to be too high for all the parameters to be calibrated simultaneously}''. Simultaneous calibration of all parameters has never been undertaken so far.

@@

When working in the Perfect Prognosis framework \citep{Klein1963}, the meteorological archive from which the predictors are extracted is usually a reanalysis dataset. Oppositely, other applications of AMs for wind forecasting are based on a Model Output Statistics context \citep[MOS, see][]{Glahn1972}, and thus use forecast archives or reforecast products \citep[e.g.][]{DelleMonache2013, DelleMonache2011, Alessandrini2015, Junk2015, Junk2015c}.

@@

The predictand (which is to be predicted) was here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss' stations network in the catchment of interest. The time series from every available gauging station were averaged over subregions of approximately 500~km$^{2}$ \citep[see][for the details]{Horton2012a} in order to smooth local effects \citep{Obled2002, Marty2012}. The time series need to be split into a calibration and an independent validation periods.

@@

Predictors for precipitation prediction, can be varied: e.g. geopotential height at different pressure levels and temporal windows, humidity variables, etc \citep[see][for a more detailed list of predictors]{BenDaoud2016}. The method often contains several levels of analogy, which are successive subsampling on predictors of different nature (eg. atmospheric circulation, moisture variables, vertical motion, and air temperature).

@@

	\item First level of analogy: $n_{1}$ dates are selected out of the preselected $n_{0}$, by means of an analogy ranking. The first level of analogy for precipitation prediction is often based on the atmospheric circulation. The similarity of the atmospheric circulation of the target date with the candidate situations is assessed on the geopotential height (at specific pressure levels, such as 500 and 1000~hPa, and over a certain spatial window, e.g. 12~h and 24~h) by means of the S1 criteria \citep[eq. (\ref{eq:S1}), ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a defined spatial window. The S1 criteria is found more relevant than Euclidean distances for pressure fields by different studies \citep{Wilson1980, Woodcock1980, Guilbaud1998, Bontron2004}.

	\begin{equation}
	\label{eq:S1}
	S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points from the grid of the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values, the more similar the pressure fields.

@@

\begin{itemize}
	\item The selection of meteorological variables (predictors) that contain synoptic scale information having direct or indirect dependency with the target predictand.
	\item The pressure level, or height, at which the predictor is selected.
	\item The temporal window is the hour(s) of the day at which the predictors are considered.
	\item The spatial window is the domain on which predictors are compared. The ideal size of this area is the one that maximizes useful information and minimizes noise.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a distance measure used for ranking observed situations according to their degree of similarity with the target situation.
	\item Eventual weights between the predictors \cite[e.g.][]{Horton2012a, Junk2015}
	\item The ideal number of analog situations, which is the best compromise in order to take into account local variability and to maximize useful synoptic information \citep{Bontron2004}.
\end{itemize}

@@

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{l} \sum_{i=1}^{l} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. The mean CRPS is averaged on the calibration, respectively the validation periods.

@@

GAs optimize a population of $N$ individuals. Each individual contains a chromosome, which is here a vector of the AM parameters. Genes are the individual parameters constituting the chromosome. They can be either categorical (e.g. choice of the meteorological variable), discrete (e.g. number of analog dates to select), or continuous.

@@

	\item Tournament selection: a number of individuals (2 or 3) are randomly picked and the best is kept, with a certain probability. This operation is performed twice, once for each partner.
\end{itemize}

@@

	\item Binary-like crossover \citep{Haupt2004}: in order to reproduce the behaviour present in the canonical crossover algorithm, which is applied to a binary representation of the genes \citep{Goldberg1989, Goldberg1990a, Herrera1998a}, \citet{Haupt2004} propose an operator that combines the standard crossover operation with an interpolation approach. The genes located after a crossover point are exchanged, but the gene located at the intersection is modified according to:
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	g_{o1,n} = g_{p1,n} - \beta (g_{p1,n} - g_{p2,n}) \\
	g_{o2,n} = g_{p2,n} + \beta (g_{p1,n} - g_{p2,n}) \\
	\end{array} \right.
	\label{eq:binary_like_crossover}
	\end{equation}
	where $g_{o1,n}$ and $g_{o2,n}$ are the $n$-th gene of the two new offspring, and $g_{p1,n}$ and $g_{p2,n}$ are those of the two parents. $\beta$ is a random value between 0 and 1, that can be unique for the whole chromosome, or that can change for every gene.

@@

The evaluated and developed mutation operators are listed hereafter (equations are only provided for the operator versions that were found relevant; the reader can refer to the corresponding literature for the details of the other options). They apply to genes made of continuous or discrete variables, but not categorical (e.g. choice of the meteorological variable or analogy criterion). In the latter case, the random choice of a new value is always based on a uniform distribution, without notion of distance in the parameters space.

@@

	\item Variable normal mutation \citep{Horton2012a}: with the same logic as the variable uniform mutation, a mutation operator was tested using a normal distribution with a variable mutation rate and standard deviation, that decrease linearly over the generations. This operator has 6 parameters.

@@

	\begin{equation}
	\varphi' = 1 - \min \left\lbrace \dfrac{G}{G_{m,r}}, 1 \right\rbrace \left(1-\omega\right)
	\end{equation}

@@

	\item Individual adaptive search radius: based on the ideas of the non-uniform mutation, we introduced a search radius in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between 0 and 1 (relatively to the parameter ranges), is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is also considered as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{eq:adaptive_search_radius_mutation}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, such as in the non-uniform mutation. No external parameter is therefore necessary.

@@

	\item Chromosome of adaptive search radius: in this operator, we combined the operations of the chromosome of adaptive mutation rate to the adaptive search radius approach. Similarly, an individual has 3 chromosomes: the first containing the values to be optimized, the second the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameter is required.

	\item Multi-scale mutation: finally, we developed another approach that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. The idea here was to test an approach that combines both exploration and exploitation during the whole optimization. Thus, the search radius $r_{a}$ of Eq. \ref{equation_mutation_rayon_adaptatif} was considered as a random value for each individual, but restricted to 4 equiprobable values: 1, 0.5, 0.1, 0.02, so a range from full exploration to fine exploitation. The only external parameter is the mutation rate which is fixed.

@@

Some constraints need to be taken into account. For example, when a crossover or a mutation operation results in a parameter value standing out of the authorized bounds, it has to be brought back within the limits. Moreover, the parameters are of different types: continuous, discrete, or categorical (e.g. the selection of the meteorological variable). New values resulting from the optimizer need to respect the type of data they represent. Other constraints exist in between the parameters, e.g. the temporal windows of the relative humidity and the precipitable water predictors have to match when processing the moisture index.

@@

In order to evaluate a combination of operators/options, 10 optimizations per configuration of GAs were processed. With all combinations of GAs options, 21,630 optimizations were performed on an Intel Xeon based HPC cluster with 96 cpu, running Linux RedHat. Such assessment was not possible on the whole archive length, and had to be performed on a reduced calibration period: 5 years, from 1998 to 2002, but looking for analog dates on a 48 years period (1961-2008). The total resources needed for this comparison amount to 229,539 h/cpu (26.2 years/cpu). One should remember that these intensive resources were needed in order to assess multiple combinations of GAs options in order to provide recommendations. Based on these, applying GAs to optimize AMs require some computational power, but not as much (the time needed is function of the archive length, the population number, GAs options, and the complexity of the AM to optimize).

@@

After a first quick assessment, GAs were found to be successful at optimizing the considered AM implementations. Figure \ref{fig:gas_evolution_good} shows the evolution of the score of the best individuals over generations for 10 optimizations processed for a given configuration of GAs (same operators options). 7 out of 10 optimizations converged to the global optimum (dashed line at the top of the plot in Fig. \ref{fig:gas_evolution_good}) after 40 to 55 generations, which is not the best performance we observed later on. The performance score of the sequential approach (bottom line) has been quickly exceeded without adding any new parameter to the method (Fig. \ref{fig:gas_evolution_good}). The improvement was about 8.8\% in this case, which is substantial.

@@

Every combination of 6 options for the couples selection (Fig. \ref{fig:operator_selectcoupl_score}) and 21 for the chromosome crossover operators (Fig. \ref{fig:operator_crossover_score}) were evaluated, along with variants of the other operators. This resulted in 1,008 combinations, requiring 10,080 optimizations.

The performances of the couples selection operators were relatively close (Fig. \ref{fig:operator_selectcoupl_score}). Overall, the tournament selection with 3~candidates was slightly superior to others, along with the roulette wheel weighting. This last one was however a bit less effective in terms of convergence and number of evaluations (not shown). The couples selection operator had no significant role in this application.

As for crossover operators (Fig. \ref{fig:operator_crossover_score}), binary-like crossover (especially with 2 points of intersection, whether $\beta$ is shared or not) was a bit better than the others, especially in terms of convergence (not shown).

@@

Having identified the leading role of the mutation operator, the next sensitivity analysis focused on it. The 10 different implementations (see section \ref{sec:gas}\ref{sec:gas:operators}\ref{sec:gas:mutation}) were tested, with different options (Fig. \ref{fig:operator_mutation_score}), bringing the number of variations up to 109. Some optimizations without any mutation were also performed as a reference. Along with variants of the other operators \citep[see][for the details]{Horton2012a}, this resulted in 660 combinations (so 6,600 optimizations).

@@

The mutation operators based on variable normal or variable uniform laws worked very poorly here and were difficult to configure. Many operators presented more or less the same performance scores and required a variable amount of assessments. The convergence analysis \citep[see][]{Horton2012a} allowed to highlight three best operators: non-uniform mutation, chromosome of adaptive search radius, and multi-scale mutation. Thus, different optimizations (an additional 4,950) were further performed using variants of these 3 operators (Fig. \ref{fig:operator_mutation_score_atmlevel}-\ref{fig:operator_mutation_score_r4}).

The first analysis was the optimization of the precipitation prediction over a subcatchment (Binn-Simplon region) in the Swiss Alps (Fig. \ref{fig:operator_mutation_score_atmlevel}). The optimizer could also choose the 2 pressure levels of the atmospheric circulation analogy (method with a single level of analogy). The resulting CRPSS performance score (section \ref{sec:am}) was obviously superior to the one obtained by the sequential calibration (bottom line on Fig. \ref{fig:operator_mutation_score_atmlevel}). For most options, it was also slightly better than the results from the optimization without selection of the pressure levels (dashed line). A clear breakthrough of the performances was not expected, as the former selection of pressure levels results from intensive comparative work \citep{Bontron2004}. This application however demonstrates that, when correctly configured, GAs can automatically and successfully choose the pressure levels. However, the automatic selection of the pressure levels significantly increased the difficulty for GAs to converge to a unique solution, ideally the global optimum. A likely explanation is that the pressure levels were considered as categorical values within the optimization (sampled with a uniform law), and thus the approaches relying on distance in the parameters space, such as the search radius, could not fully exploit the properties that made them efficient. However, even though the results showed a certain variability, most of them presented very good performance scores, despite different selection of AM parameters.

The same experiment was performed for another region (Swiss Chablais, see Fig. \ref{fig:map}), which is sensitive to other meteorological influences (Fig. \ref{fig:operator_mutation_score_rhoneamont}), in order to assess eventual dependencies of the operators with the predictand. Even though differences can be observed in Fig. \ref{fig:operator_mutation_score_atmlevel}, it is globally the same options that perform better.

Next, a second level of analogy was introduced (Fig. \ref{fig:operator_mutation_score_r2}) based on moisture variables (section \ref{sec:am}). GAs had to optimize both levels of analogy (geopotential height and moisture index) simultaneously. Once again, the results were better than the sequential calibration (bottom line on Fig. \ref{fig:operator_mutation_score_r2}). And finally, a preselection on air temperature was added instead of the fixed calendar window, as proposed by \cite{BenDaoud2016}. The results showed generally higher scores (Fig. \ref{fig:operator_mutation_score_r4}), demonstrating the success of the optimizer to take advantage of this new degree of freedom, and its ability to handle optimization of 3 analogy levels jointly. Again, the most relevant options were globally the same.

>>>>

@@

\abstract{Analog methods are based on a statistical relationship between synoptic meteorological variables (predictors) and local weather (predictand, which are to be predicted), which is defined by several parameters that are often calibrated by means of a semi-automatic sequential procedure. This calibration approach is fast, but has strong limitations: it is made of successive steps and thus cannot handle all parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows (hour(s) of the day) on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely Genetic Algorithms, which can optimize jointly all parameters of the method and can get closer to a global optimum, by taking into account the parameters dependencies. Moreover, it can objectively calibrate parameters that were previously manually assessed, and can take into account new degrees of freedom.
Genetic Algorithms need however to be tailored to the addressed problem. Multiple combinations of algorithms were assessed, and new ones were developed (e.g., the \textit{chromosome of adaptive search radius} which was found to be very robust), in order to make recommendations for the use of Genetic Algorithms to optimize several variants of analog methods. A global optimization approach opens new perspectives for the improvement of analog methods, and their application to new regions or to new predictands.
}

@@

Analog methods (AMs) rely on the hypothesis that similar situations in terms of the atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. They consist in sampling a certain number of past situations based on different synoptic-scale meteorological variables (predictors) in order to build a probabilistic prediction for a local weather variable of interest (predictand). A common usage of AMs is for operational precipitation forecasting \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015b, BenDaoud2016}, or more recently for precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. However, AMs or equivalent are also used to predict temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

@@

A common approach to optimize the method is by means of a semi-automatic sequential calibration procedure, developed by \citet{Bontron2004}, also described in \citet{BenDaoud2016}, and extended by \citet{Radanovics2013}. It determines some parameters of the method, sequentially for each consecutive analogy level (e.g., on the atmospheric circulation or on a moisture index). It starts by a manual selection of the meteorological variables (e.g., geopotential height, relative humidity, etc.), the pressure levels, the temporal windows, and the initial analog numbers. Then, the spatial window on which the predictors are compared is optimized by an iterative growth of the domain, and the number of analogs is finally reassessed. A successive level of analogy can then be added and its spatial window optimized. Parameters of the preceding levels of analogy are not reassessed, except for the number of analog situations to keep.

@@

When creating the sequential calibration procedure, \citet{Bontron2004} was aware of the problem of dependencies between parameters and wrote: ''\textit{We perceive here the combinatorial aspect of our problem: variables and spatial windows are not independent. We will present our results by first searching the best variable (e.g., selection of the pressure level and the temporal window for the geopotential height) on a chosen spatial window, and next, the best window for the chosen variable. However, even by repeating the process, are we sure to obtain the optimal combination?}''. And later in his work: ''\textit{Our approach, which is again to vary the parameters one by one -- the others being fixed in a more or less arbitrary manner -- may therefore not exactly lead us to the optimal solution}''. \citet{Bliefernicht2010} has also faced the combinatorial issue of AMs parameters and concluded that one needs to be an expert to have a sense of their respective influence, sensitivity and nonlinear interactions. \citet{BenDaoud2010}, when calibrating an AM, also stated that ''\textit{the combinatory aspect related to the calibration was found to be too high for all the parameters to be calibrated simultaneously}''. Simultaneous calibration of all parameters has never been undertaken so far.

@@

When working in the Perfect Prognosis framework \citep{Klein1963}, the meteorological archive from which the predictors are extracted is usually a reanalysis dataset. Oppositely, other applications of AMs for wind forecasting are based on a Model Output Statistics context \citep[MOS, see][]{Glahn1972}, and thus use forecast archives or reforecast products \citep[e.g.,][]{DelleMonache2013, DelleMonache2011, Alessandrini2015, Junk2015, Junk2015c}.

@@

The predictand (which is to be predicted) was here the daily precipitation (0540~UTC to 0540~UTC the following day) measured at the MeteoSwiss' stations network in the catchment of interest. The time series from every available gauging station were averaged over subregions of approximately 500~km$^{2}$ \citep[see][for the details]{Horton2012a} in order to smooth local effects \citep{Obled2002, Marty2012}. The time series need to be split into a calibration and an independent validation periods.

@@

Predictors for precipitation prediction, can be varied: e.g., geopotential height at different pressure levels and temporal windows, humidity variables, etc \citep[see][for a more detailed list of predictors]{BenDaoud2016}. The method often contains several levels of analogy, which are successive subsampling on predictors of different nature (e.g., atmospheric circulation, moisture variables, vertical motion, and air temperature).

@@

	\item First level of analogy: $n_{1}$ dates are selected out of the preselected $n_{0}$, by means of an analogy ranking. The first level of analogy for precipitation prediction is often based on the atmospheric circulation. The similarity of the atmospheric circulation of the target date with the candidate situations is assessed on the geopotential height (at specific pressure levels, such as 500 and 1000~hPa, and over a certain spatial window, e.g., 12~h and 24~h) by means of the S1 criteria \citep[Eq. (\ref{eq:S1}), ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a defined spatial window. The S1 criteria is found more relevant than Euclidean distances for pressure fields by different studies \citep{Wilson1980, Woodcock1980, Guilbaud1998, Bontron2004}.

	\begin{equation}
	\label{eq:S1}
	S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
	{\displaystyle \sum_{i} max( \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert ) }
	\end{equation}
	where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points from the grid of the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values, the more similar the pressure fields.

@@

\begin{itemize}
	\item The selection of meteorological variables (predictors) that contain synoptic scale information having direct or indirect dependency with the target predictand.
	\item The pressure level, or height, at which the predictor is selected.
	\item The temporal window is the hour(s) of the day at which the predictors are considered.
	\item The spatial window is the domain on which predictors are compared. The ideal size of this area is the one that maximizes useful information and minimizes noise.
	\item The analogy criterion, needed to compare the variables on the chosen spatial and temporal windows, is a distance measure used for ranking observed situations according to their degree of similarity with the target situation.
	\item Eventual weights between the predictors \cite[e.g.,][]{Horton2012a, Junk2015}
	\item The ideal number of analog situations, which is the best compromise in order to take into account local variability and to maximize useful synoptic information \citep{Bontron2004}.
\end{itemize}

@@

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{l} \sum_{i=1}^{l} \left\{  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right\} 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value one otherwise. The mean CRPS is averaged on the calibration, respectively the validation periods.

@@

GAs optimize a population of $N$ individuals. Each individual contains a chromosome, which is here a vector of the AM parameters. Genes are the individual parameters constituting the chromosome. They can be either categorical (e.g., choice of the meteorological variable), discrete (e.g., number of analog dates to select), or continuous.

@@

	\item Tournament selection: a number of individuals (two or three) are randomly picked and the best is kept, with a certain probability. This operation is performed twice, once for each partner.
\end{itemize}

@@

	\item Binary-like crossover \citep{Haupt2004}: in order to reproduce the behaviour present in the canonical crossover algorithm, which is applied to a binary representation of the genes \citep{Goldberg1989, Goldberg1990a, Herrera1998a}, \citet{Haupt2004} propose an operator that combines the standard crossover operation with an interpolation approach. The genes located after a crossover point are exchanged, but the gene located at the intersection is modified according to:
	\begin{equation}
	\left\lbrace \begin{array}{l} 
	g_{o1,n} = g_{p1,n} - \beta (g_{p1,n} - g_{p2,n}) \\
	g_{o2,n} = g_{p2,n} + \beta (g_{p1,n} - g_{p2,n}) \\
	\end{array} \right.
	\label{eq:binary_like_crossover}
	\end{equation}
	where $g_{o1,n}$ and $g_{o2,n}$ are the $n$-th gene of the two new offspring, and $g_{p1,n}$ and $g_{p2,n}$ are those of the two parents. $\beta$ is a random value between zero and one, that can be unique for the whole chromosome, or that can change for every gene.

@@

The evaluated and developed mutation operators are listed hereafter (equations are only provided for the operator versions that were found relevant; the reader can refer to the corresponding literature for the details of the other options). They apply to genes made of continuous or discrete variables, but not categorical (e.g., choice of the meteorological variable or analogy criterion). In the latter case, the random choice of a new value is always based on a uniform distribution, without notion of distance in the parameters space.

@@

	\item Variable normal mutation \citep{Horton2012a}: with the same logic as the variable uniform mutation, a mutation operator was tested using a normal distribution with a variable mutation rate and standard deviation, that decrease linearly over the generations. This operator has six parameters.

@@

	\begin{equation}
	\varphi' = 1 - \min \left( \dfrac{G}{G_{m,r}}, 1 \right) \left(1-\omega\right)
	\end{equation}

@@

	\item Individual adaptive search radius: based on the ideas of the non-uniform mutation, we introduced a search radius in the approach of individual adaptive mutation rates. This search radius $r_{a}$, bounded between zero and one (relatively to the parameter ranges), is also adaptive and behaves similarly to the adaptive mutation rates. In order to separate its evolution from the one of the mutation rate, its own value is also considered as a self-mutation rate to eventually mutate before being used as a normalized search radius. The value of a mutated gene is given by the following equation, which is a simplification of the non-uniform mutation:
	\begin{equation}
	g_{n}^{'} = 
	\left\lbrace \begin{array}{l l} 
	g_{n} + \left(b_{n}-g_{n}\right) r_{2} r_{a} & if \; r_{1} < 0.5 \\
	g_{n} - \left(g_{n}-a_{n}\right) r_{2} r_{a} & if \; r_{1} \geq 0.5 \\
	\end{array} \right.
	\label{eq:adaptive_search_radius_mutation}
	\end{equation}
	where $r_{1}$ and $r_{2}$ are randomly selected, such as in the non-uniform mutation. No external parameter is therefore necessary.

@@

	\item Chromosome of adaptive search radius: in this operator, we combined the operations of the chromosome of adaptive mutation rate to the adaptive search radius approach. Similarly, an individual has three chromosomes: the first containing the values to be optimized, the second the distributed mutation rate, and the last one, the distributed search radius. Again, no external parameter is required.

	\item Multi-scale mutation: finally, we developed another approach that is also based on the search radius concept. However, the latter is not decreasing with time. Methods based on a reduction of the mutation rate or radius simulate a transition from the exploration phase to the exploitation one. The idea here was to test an approach that combines both exploration and exploitation during the whole optimization. Thus, the search radius $r_{a}$ of Eq. (\ref{eq:adaptive_search_radius_mutation}) was considered as a random value for each individual, but restricted to four equiprobable values: 1, 0.5, 0.1, 0.02, so a range from full exploration to fine exploitation. The only external parameter is the mutation rate which is fixed.

@@

Some constraints need to be taken into account. For example, when a crossover or a mutation operation results in a parameter value standing out of the authorized bounds, it has to be brought back within the limits. Moreover, the parameters are of different types: continuous, discrete, or categorical (e.g., the selection of the meteorological variable). New values resulting from the optimizer need to respect the type of data they represent. Other constraints exist in between the parameters, e.g., the temporal windows of the relative humidity and the precipitable water predictors have to match when processing the moisture index.

@@

In order to evaluate a combination of operators/options, 10 optimizations per configuration of GAs were processed. With all combinations of GAs options, 21~630 optimizations were performed on an Intel Xeon based HPC cluster with 96 cpu, running Linux RedHat. Such assessment was not possible on the whole archive length, and had to be performed on a reduced calibration period: five years, from 1998 to 2002, but looking for analog dates on a 48 years period (1961-2008). The total resources needed for this comparison amount to 229~539 h cpu (26.2 yr cpu). One should remember that these intensive resources were needed in order to assess multiple combinations of GAs options in order to provide recommendations. Based on these, applying GAs to optimize AMs require some computational power, but not as much (the time needed is function of the archive length, the population number, GAs options, and the complexity of the AM to optimize).

@@

After a first quick assessment, GAs were found to be successful at optimizing the considered AM implementations. Figure \ref{fig:gas_evolution_good} shows the evolution of the score of the best individuals over generations for 10 optimizations processed for a given configuration of GAs (same operators options). Seven out of 10 optimizations converged to the global optimum (dashed line at the top of the plot in Fig. \ref{fig:gas_evolution_good}) after 40 to 55 generations, which is not the best performance we observed later on. The performance score of the sequential approach (bottom line) has been quickly exceeded without adding any new parameter to the method (Fig. \ref{fig:gas_evolution_good}). The improvement was about 8.8\% in this case, which is substantial.

@@

Every combination of six options for the couples selection (Fig. \ref{fig:operator_selectcoupl_score}) and 21 for the chromosome crossover operators (Fig. \ref{fig:operator_crossover_score}) were evaluated, along with variants of the other operators. This resulted in 1~008 combinations, requiring 10~080 optimizations.

The performances of the couples selection operators were relatively close (Fig. \ref{fig:operator_selectcoupl_score}). Overall, the tournament selection with three candidates was slightly superior to others, along with the roulette wheel weighting. This last one was however a bit less effective in terms of convergence and number of evaluations (not shown). The couples selection operator had no significant role in this application.

As for crossover operators (Fig. \ref{fig:operator_crossover_score}), binary-like crossover (especially with two points of intersection, whether $\beta$ is shared or not) was a bit better than the others, especially in terms of convergence (not shown).

@@

Having identified the leading role of the mutation operator, the next sensitivity analysis focused on it. The 10 different implementations (see section \ref{sec:gas}\ref{sec:gas:operators}\ref{sec:gas:mutation}) were tested, with different options (Fig. \ref{fig:operator_mutation_score}), bringing the number of variations up to 109. Some optimizations without any mutation were also performed as a reference. Along with variants of the other operators \citep[see][for the details]{Horton2012a}, this resulted in 660 combinations (so 6~600 optimizations).

@@

The mutation operators based on variable normal or variable uniform laws worked very poorly here and were difficult to configure. Many operators presented more or less the same performance scores and required a variable amount of assessments. The convergence analysis \citep[see][]{Horton2012a} allowed to highlight three best operators: non-uniform mutation, chromosome of adaptive search radius, and multi-scale mutation. Thus, different optimizations (an additional 4~950) were further performed using variants of these three operators (Fig. \ref{fig:operator_mutation_score_atmlevel}-\ref{fig:operator_mutation_score_r4}).

The first analysis was the optimization of the precipitation prediction over a subcatchment (Binn-Simplon region) in the Swiss Alps (Fig. \ref{fig:operator_mutation_score_atmlevel}). The optimizer could also choose the two pressure levels of the atmospheric circulation analogy (method with a single level of analogy). The resulting CRPSS performance score (section \ref{sec:am}) was obviously superior to the one obtained by the sequential calibration (bottom line on Fig. \ref{fig:operator_mutation_score_atmlevel}). For most options, it was also slightly better than the results from the optimization without selection of the pressure levels (dashed line). A clear breakthrough of the performances was not expected, as the former selection of pressure levels results from intensive comparative work \citep{Bontron2004}. This application however demonstrates that, when correctly configured, GAs can automatically and successfully choose the pressure levels. However, the automatic selection of the pressure levels significantly increased the difficulty for GAs to converge to a unique solution, ideally the global optimum. A likely explanation is that the pressure levels were considered as categorical values within the optimization (sampled with a uniform law), and thus the approaches relying on distance in the parameters space, such as the search radius, could not fully exploit the properties that made them efficient. However, even though the results showed a certain variability, most of them presented very good performance scores, despite different selection of AM parameters.

The same experiment was performed for another region (Swiss Chablais, see Fig. \ref{fig:map}), which is sensitive to other meteorological influences (Fig. \ref{fig:operator_mutation_score_rhoneamont}), in order to assess eventual dependencies of the operators with the predictand. Even though differences can be observed with Fig. \ref{fig:operator_mutation_score_atmlevel}, it is globally the same options that perform better.

Next, a second level of analogy was introduced (Fig. \ref{fig:operator_mutation_score_r2}) based on moisture variables (section \ref{sec:am}). GAs had to optimize both levels of analogy (geopotential height and moisture index) simultaneously. Once again, the results were better than the sequential calibration (bottom line on Fig. \ref{fig:operator_mutation_score_r2}). And finally, a preselection on air temperature was added instead of the fixed calendar window, as proposed by \cite{BenDaoud2016}. The results showed generally higher scores (Fig. \ref{fig:operator_mutation_score_r4}), demonstrating the success of the optimizer to take advantage of this new degree of freedom, and its ability to handle optimization of three analogy levels jointly. Again, the most relevant options were globally the same.
